---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### Unzip the file

The original repository contains a zip file of the data, as shown below:
```{r}
dir()
```
We first unzip the file and examine the contents again:
```{r}
unzip("activity.zip")
dir()
```

### Reading the data

```{r}
library(tidyverse)
d <- as_tibble(read_csv("activity.csv"))
d
```
We see that the data are read in successfully.

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- d %>% group_by(date) %>% dplyr::summarize(steps=sum(steps))
```
Plot a histogram:
```{r}
hist(steps_per_day$steps, xlab="Steps")
```
We see that the person usually walks 10000 to 15000 steps per day.

* Mean steps per day: `r format(mean(steps_per_day$steps, na.rm=T), 2)`
* Median steps per day: `r format(median(steps_per_day$steps, na.rm=T), 2)`

## What is the average daily activity pattern?

To answer this question we need to look at all the intervals.
```{r}
mean_steps <- d %>% group_by(interval) %>% dplyr::summarize(mean=mean(steps, na.rm=T))
```
Then make a line plot to look at the activity pattern:
```{r}
plot(mean_steps$interval, mean_steps$mean, type="l", xlab="Interval", ylab="Number of steps", main="Average number of steps per interval")
ix <- which.max(mean_steps$mean) #index of max steps
```
The interval with the maximum number of steps is `r mean_steps$interval[ix]`.

## Imputing missing values

Calculate the total number of missing values:
```{r}
sum(is.na(d$steps))
```
Impute missing values with the mean, using the impute function from the Hmisc package:
```{r}
library(Hmisc)
d$steps <- impute(d$steps, fun=mean)
```
Below we repeat the histogram from above to check the effect of imputation.
```{r}
steps_per_day <- d %>% group_by(date) %>% dplyr::summarize(steps=sum(steps))
```
Plot a histogram:
```{r}
hist(steps_per_day$steps, xlab="Steps")
```

* Mean steps per day: `r format(mean(steps_per_day$steps, na.rm=T), 2)`
* Median steps per day: `r format(median(steps_per_day$steps, na.rm=T), 2)`

We can see some effects of the imputation. A more sophisticated imputation strategy would be better.

## Are there differences in activity patterns between weekdays and weekends?

First we create the day type factor.
```{r}
d$day_type <- as.factor(ifelse(weekdays(d$date)=="Saturday"|weekdays(d$date)=="Sunday", "weekend", "weekday"))
```
Then plot the steps for each day type:
```{r}
library(lattice)
mean_steps <- d %>% group_by(interval, day_type) %>% dplyr::summarize(mean=mean(steps, na.rm=T))
xyplot(mean_steps$mean ~ mean_steps$interval | day_type, data=mean_steps, type="l", ylab = "Number of steps", main="Activity pattern for weekdays and weekends", xlab = "Interval", layout=c(1,2))
```


